### 2014-10-01 paper/include/background.tex
bg work

\begin{figure*}[t]
    \centering
    \subfigure[DRAM hierarcy]{
        \centering
        \includegraphics[width=0.4\textwidth]{dram_orga}
		\label{fig:dram_orga}
    }
    \subfigure[DRAM bank structure]{
        \centering
        \includegraphics[width=0.4\textwidth]{dram_orgb}
		\label{fig:dram_orgb}
    }%
	\caption{DRAM system organization \cite{raidr}.}
	\label{fig:dram_org}
\end{figure*}

@@

The memory controller (MC) normally issues the refresh operation, the auto-refresh command, periodically at an interval of 64 ms for each row, an interval that has been constant for the latests DRAM specifications. The interval is halved if the DRAM exeeds the specified working temprature of 85$^{\circ}$C.

The simplest approach to refresh one rank is to refresh all banks in parallel row-by-row at the start of the 64ms interval. This approch is called Burst refresh and has some downsides. The approch blocks normal accesses and form a peak of power consumption during the burst time. A better alternative is to distribute the refreshes across the interval, resulting in a lower average latency for the processor and a more stable power consumption.

Explain different refresh modes: Burst refresh vs distributed refresh and RAS-only refresh vs CAS before RAS refresh.

>>>>

\begin{figure*}[t]
    \centering
    \subfigure[DRAM hierarcy]{
        \centering
        \includegraphics[width=0.4\textwidth]{dram_orga}
		\label{fig:dram_orga}
    }
    \subfigure[DRAM bank structure]{
        \centering
        \includegraphics[width=0.4\textwidth]{dram_orgb}
		\label{fig:dram_orgb}
    }
	\caption{DRAM system organization \cite{raidr}.}
	\label{fig:dram_org}
\end{figure*}

@@

The memory controller (MC) normally issues the refresh operation, the auto-refresh command, periodically at an interval of 64 ms for each row, an interval that has been constant for the latests DRAM specifications. The interval is halved if the DRAM exeeds the specified working temprature of 85$^{\circ}$C. The average time between refresh commands are issued by the MC is denoted \(t_{REFI}\) and the time it takes to perform the refresh is denoted \(t_{RFC}\), i.e the time where no normal access requests can be serviced.

To refresh one rank, the simplest approach is to refresh all banks in parallel row-by-row at the start of the 64ms interval. This approch is called Burst refresh and has some downsides. The approch blocks normal accesses for a longer period and form a peak of power consumption during the burst time. A better alternative is to distribute the refreshes across the interval, resulting in a lower average latency for the processor and a more stable power consumption.

The implementation of a DRAM refresh cycle can be made in two different ways, to either affect one row or many. The first is called RAS-only refresh (ROR) where refresh cycle refreshes one targeted row. The address for the targeted row has to be put on the bus and it is the MC that makes sure that a row are refreshed within the 64ms interval. The second is called CAS before RAS refresh (CBR). In this scheme the memory module has an internal address counter that displays which row that shall be refreshed. On a refresh cycle, the row stated by the internal address counter is refreshed and the counter is then incremented. The advantage with this method is that the address is never put on the bus, which gives a lower power consumption and less bus usage than ROR. CBR is the most used method and is the one normally implemented in the auto-refresh command.

When using auto-refresh (CBR), all banks in a rank are blocked for normal accesses. All banks are also blocked when ROR is used due to DRAM chip design, however some DRAM chips support per-bank refresh commands that makes it possible to only block the specific bank which is getting refreshed.

As m

\begin{figure}[t]
    \centering
    \subfigure[DRAM throughput loss]{
        \centering
        \includegraphics[width=0.33\textwidth]{dram_throughput}
        \label{fig:dram_throughput}
    }
    \subfigure[DRAM power consumption]{
        \centering
        \includegraphics[width=0.33\textwidth]{dram_power}
        \label{fig:dram_power}
    }%
    \caption{Projections on DRAM throughput loss and power consumption \cite{raidr}.}
    \label{fig:dram_data_proj}
\end{figure}
